---
title: "Lab 7: Visualizing International Space Station Information via APIs"
author: "Allen Choi and Hannah Pawig"
format:
  html:
    code-fold: show
    code-summary: "Click to hide/show the code"
    code-overflow: wrap
    toc: true
embed-resources: true
editor: source
---


```{r}
#| message: false
#| echo: false
#| include: false
#| label: loading-packages

# Load the required libraries
library(jsonlite)
library(httr)
library(tidyverse)
library(leaflet)
```



### Pass Times for U.S. State Captials

## 1.
Use the API to get the next **three** predicted pass times for all of the 
U.S. state capitals. Organize all of this information into a nice data frame (or
data frame-like object) that will be easy to use for mapping.


```{r}
#| label: US-state-capital-data
#| echo: false
#| message: false
#| include: false
# Read in text file of US state capitals' lat and long
state_ll <- read_table(
  file = "https://people.sc.fsu.edu/~jburkardt/datasets/states/state_capitals_ll.txt",
  col_names = c("state", "lat", "long")) |> 
  filter(state != "US") |>  # Remove the row with China
  filter(state != "AK") # get rid of alaska since ISS doesn't orbit it

state_names <- read_delim(
  file = "https://people.sc.fsu.edu/~jburkardt/datasets/states/state_capitals_name.txt",
  delim = "  ",
  col_names = c("state", "name")
)
```

- We realized that the state_capitals_ll.txt file has a pair of coordinates in China, so we filtered that out.
- We also filtered out the Alaska capital (Juneau) because the ISS doesn't orbit over Alaska.


- Below we have the API request functions `get_pass_times` and `create_passtimes_df`. The first function is a helper function that makes one GET() request from the provided Pass Times API found on this [page]((https://g7vrd.co.uk/public-satellite-pass-rest-api).
- The second function takes in a data frame, the latitude and longitude columns of that dataframe, and optionally the minimum elevation and hours lapsed. The `create_passtimes_df` function will loop through the dataframe and call the `get_pass_times` function for each row, storing the results in a new dataframe. The new dataframe will have columns for the state, latitude, longitude, and the three pass times for every capital in the provided dataframe (each row).
```{r}
#| label: API-request-fn
# Need a query for all latitude and longitudes in state_ll
# and get the next three passes for each
# Create a function to get the next three passes for each state capital in state_ll
# Helper function to get pass times
get_pass_times <- function(lat, long, minelevation = 30, hours = 48) {
  url <- paste0("https://api.g7vrd.co.uk/v1/satellite-passes/25544/", lat,"/",
                long,".json?minelevation=", minelevation,"&","hours=", hours)
  res <- GET(url)
  
  # Check if the request was successful
  if (status_code(res) == 200) {
    passtimes <- res$content |>
      rawToChar() |>
      fromJSON() |>
      data.frame() |>
      select(passes.tca)
    return(t(passtimes$passes.tca))
  } else {
    stop("Error: Unable to retrieve data from the API. Please check the URL or parameters.")
  }
}

```


```{r}
# Function to pass in a dataframe to get all times the satellite passes
#| label: get-all-passes-fn

create_passtimes_df <- function(df, state_col, lat_col, long_col, minelevation = 70, hours = 12) {
  
  # store the columns in respective variables
  state <- df |> select({{state_col}})
  lat <- df |> select({{lat_col}})
  long <- df |> select({{long_col}})
  
  # Create the base dataframe, with columns for the 3 passes we'll get
  return_df <- data.frame(state = state,
                          lat = lat,
                          long = long,
                          time1 = NA,
                          time2 = NA,
                          time3 = NA)
  # Loop through each row of the dataframe to request the pass times one row at a time
  for (i in 1:nrow(return_df)) {
    # Call the function to get the pass times
    cur_pass_times <- get_pass_times(lat = return_df[i, 2], # get the latitude for row i
                                     long = return_df[i, 3], # get longitude for row i
                                     minelevation, hours)
    
    # Add the pass times to the dataframe with tidyr
    # and select the first three times
    
    # sometimes less than 3 times are available, need to handle that
    if (length(cur_pass_times) == 1) {
      cur_pass_times <- c(cur_pass_times[, 1], NA, NA)
    } else if (length(cur_pass_times) == 2) {
      cur_pass_times <- c(cur_pass_times[, c(1, 2)], NA)
    } else {
      cur_pass_times <- cur_pass_times[,c(1,2,3)]
    }
    
    return_df[i, c(4, 5, 6)] <- cur_pass_times
            
    
    Sys.sleep(1) # get sys sleep so API doesn't get angry
  }
  
  return(return_df)
}

```

```{r}
#| label: create-tidy-df
final_dataset <- create_passtimes_df(state_ll, state, lat, long, minelevation = 30, hours = 72)
final_dataset <- final_dataset |> 
  mutate(across(time1:time3, ~as_datetime(.x))) |>  # convert passtimes to datetimes
  arrange(time1) |>  # sort by first passes
  left_join(state_names, by = join_by(state == state))
```






### Mapping the Data, Drawing the Route of the ISS

## 2.
Map the U.S. state capitals using `leaflet`.


## 3.
Find an image of your choosing to use as the marker icon, that's relevant 
for these data. Maybe a spaceship? 🚀 Or an astronaut? Or the moon?

## 4. 
The hover-over information for each marker should include the name of the
state capital and the soonest predicted pass time. The click information should
include the name of the state capital and all three predicted pass times. Make
sure these times are in an easy-to-read format.


## 5.
Overlay lines on your map that connect the U.S. state capitals in order of
pass times (essentially east to west) so that we can see the expected pass
order.

- We combined questions 2-5 in the following code chunk.
```{r}
#| label: create-leaflet-map

# icon: https://www.flaticon.com/free-icons/moon
moon_icon <- makeIcon(
  iconUrl = "https://cdn-icons-png.flaticon.com/128/616/616465.png",
  iconWidth = 20,
  iconHeight = 20
)

final_dataset |> 
  group_by(state) |> # group times by state 
  leaflet() |> 
  addTiles() |> 
  addPolylines(
    lng = final_dataset$long,
    lat = final_dataset$lat,
    color = "#2ca25f",
    opacity = 0.7,
    weight = 5
  ) |> 
  addMarkers(
    lng = final_dataset$long,
    lat = final_dataset$lat,
    icon = moon_icon,
    popup = paste0("State Capitol Name: ", " ", final_dataset$name, " <br> ",
                   "All three times: ", " <br> ",
                   "First Time: <br> ",
                   final_dataset$time1, " <br> ",
                   "Second Time: <br> ",
                   final_dataset$time2, " <br> ",
                   "Third Time: <br> ",
                   final_dataset$time3),
    label =  paste0("State Capitol Name: ", " ", final_dataset$name,
                   "; Soonest possible time: ", final_dataset$time1)
  )

```

```{r}

final_dataset |> 
  head(1) |> 
  select(name, state)
```

- The predicted first pass in the United States happens over Sacramento, CA (at the time this was made).
- The polylines are plotted in order of the pass times, which was made possible because
we sorted the `final_dataset` by the first pass time of each capital.